Cluster Difference Imaging Photometric Survey. III. An Open Cluster With a 250\,pc Halo

### 2021-02-02 cluster_paper/ms.tex
paper updates

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations (``halos'') surrounding nearby open clusters.  The stars in these
  halos could have escaped from the cores of the cluster, or they
  could be remnants of a star formation
  complex that has since dispersed.  They could also be field
  stars---the false positives of overly aggressive clustering algorithms.
  In this study, we focus on a single reported 250\,pc-long halo
  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical tidal radius of
  this cluster is $\sim$XX\,pc.  Combining photometry from Gaia,
  rotation periods from TESS, and lithium measurements from
  the Gaia-ESO and GALAH surveys, we find that the halo of
  NGC\,2516 is real and coeval with the core. At fixed stellar
  mass, the rotation rate and
  lithium abundance are strongly anti-correlated. And finally, given Gaia-selected halo
  members, the contamination rate of field stars in the halo (XX\%) is
  larger than that in the core (YY\%).  This work expands the set of
  confirmed NGC\,2516 members by a factor of $\approx$2, and
  quantifies the degree to which Gaia-based analyses can 
  identify dispersed stellar populations against a background of field
  star contaminants.  We highlight the implications for spectroscopic
  survey targeting, open cluster dispersal, and planet searches around
  young stars.
\end{abstract}

@@

@@

		\subfloat{
			\includegraphics[width=0.75\textwidth]{f8b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
    velocity space, cross-matched against the rotators, but now in
    physical units.}
    Top: relevant projections. Bottom: binned versus 3d position and
    2d velocity difference.
		\label{fig:physical_x_rotn}
	}
\end{figure*}

@@

Figure~\ref{fig:gaia6d_x_rotn} shows the result.  An important nuance
in interpreting this plot is that we show only stars with
$0.5<(Bp-Rp)_0<1.2$ for which our TESS pipeline successfully made
detrended light curves.  In other words---the base ``core'' and
``halo'' samples are the stars for which we could have plausibly
detected a rotation period.
Faint stars with $Rp>16$ were beyond our light curve selection
limit, and crowded stars ({\it e.g.,} near saturated stars
in the cluster center) do not appear.
We selected the color limit above based on Figure~\ref{fig:rot},
since the density of period detections seems to decrease once 
$(Bp-Rp)_0 \gtrsim 1.25$, {\it i.e.}, for spectral types later
than $\approx$K4V.

>>>>

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  The
  stars in these ``halos'' could have escaped from the cores of the
  cluster, or they could be remnants of a star formation complex that
  has since dispersed.  They could also be field stars---false
  positives from overly aggressive clustering algorithms.  In this
  study, we focus on a 250\,pc halo reported
  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical
  tidal radius of the cluster is {\bf $\sim$XX\,pc}.  Using photometry
  from Gaia, rotation periods from TESS, and lithium measurements from
  Gaia-ESO and GALAH, we find that the halo of NGC\,2516 is real and
  coeval with the core.  Two in three kinematically-selected halo
  members 25-250\,pc from the cluster center show rotation periods
  consistent with a gyrochronological age of 120\,Myr.  Stars in the
  inner 25\,pc show higher period detection rates than the
  halo (73\% vs.\ 66\%), but both vastly exceed our field star
  comparison sample ({\bf XX\%}).  The rotation rate and lithium abundance
  are also anti-correlated at fixed stellar mass, as is the case with
  comparable clusters.  This work expands the set of confirmed
  NGC\,2516 members by a factor of $\approx$2, and quantifies the
  degree to which Gaia-based analyses can identify dispersed stellar
  populations against a background of field star contaminants.  We
  highlight the implications for spectroscopic survey targeting, open
  cluster dispersal, and planet searches around young stars.
\end{abstract}
% Rotators < 25pc : 0.72929 +/- 0.00013
% Rotators > 25pc : 0.66172 +/- 0.00023

@@

{\bf Emphasize that figure 1 does not clearly tell us whether or not the halo is really distinct from the field, by eye.}

@@

		\subfloat{
			\includegraphics[width=0.75\textwidth]{f8b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
    velocity space, cross-matched against the rotators, but now in
    physical units.}
    Top: relevant projections. Bottom: binned versus 3d position and
    2d velocity difference.
    Bins have width 25\,pc in position, and 2.0\,km\,s$^{-1}$ in
    velocity.
		\label{fig:physical_x_rotn}
	}
\end{figure*}

@@

Figure~\ref{fig:gaia6d_x_rotn} shows the result.  An important nuance
in interpreting this plot is that we show only stars with
$0.5<(Bp-Rp)_0<1.2$ for which our TESS pipeline successfully made
detrended light curves.  In other words---the base sample is the stars
for which we could have plausibly detected a rotation period.  Faint
stars with $Rp>16$ were beyond our light curve selection limit, and
crowded stars ({\it e.g.,} near saturated stars in the cluster center)
do not appear.  We selected the color limit above based on
Figure~\ref{fig:rot}, since the density of period detections seems to
decrease once $(Bp-Rp)_0 \gtrsim 1.25$, {\it i.e.}, for spectral types
later than $\approx$K4V.
Our expectation is that completion effects become important for
fainter stars.
At the distance of NGC\,2516, a K4V star has $T\approx 14.5$,
for which our reductions of the TESS images yield a precision of
$\approx 7$\,mmag per half-hour exposure \citep{bouma_wasp4b_2019}.
{\bf 
The AMPLITUDES of the rotation signals are like 10-100 mmag (CITE
REBULL 2020, FIGURE 10).
}

%G=18 is like Abs G 9.8, Bp-Rp not corrected of 2.7. 
% (Bp-Rp )_0 of 1.25 is (Bp-Rp) of 1.38... or Abs G of ~= 6.8.
% or G-Rp ~= 0.7
% So, G~=15... meaning Rp~=14.3.

% IN POSITION
% Bin width: 25
% Rotators: [361  82  43  30  16  18  14   8   6   2   4   0   0   0   0
% 0   0   0 0   0]
% Comparison: [495 116  65  46  28  20  22  12  11   8   5   1   0   1
% 0   2   0   0 0   0]
% Rotators < 25pc : 0.72929 +/- 0.00013
% Rotators > 25pc : 0.66172 +/- 0.00023
% ... where numerator and denom are 223/337
